---
title: "Visualizing July Lipid Data" #Goal: determine the explanatory variable for the accumulation of lipid mass given two different strains and given larvae exposed to two treatments.
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r}
##loading libraies
library(tidyr)
library(ggplot2)
library(plyr)
library(dplyr)
library(MASS)
library(readxl)
library(doBy)
```

````{r Reading in the Data, removing NAs, and making new columns to separate possible explanatory variables}
## reading in the data and looking at its structure
setwd("/Users/JamesB/Google Drive/Graduate School/Self_JamesTBrown/GitHub/Lipid_Quantification/Scripts_Analysis")
data=read_excel("/Users/JamesB/Google Drive/Graduate School/Self_JamesTBrown/GitHub/Lipid_Quantification/Data/20171207 Samples RAW data.xlsx", sheet = "2017JUL")

#converting wide to long. This code puts all the observations into one column with the label of your choice (measurements used below)
dat.long=gather(data,traits,measurement,wet_mass:lipid_mass)

##This code breaks apart the data in the column "treat" into "strain" and "photoperiod"
data$strain<-substr(data$treat,1,2)
data$photoperiod<-as.numeric(substr(data$treat,3,4))

##Exclude NAs from the data set
data2<-na.omit(data)

##Setting Photoperiod as a numeric variable so it can be quantified by the model
data2$photoperiod<-as.factor(as.character(data2$photoperiod))

##Adding an SD column to data set


str(data2) #shows the structure of the data
````


```{r model construction}
## Intrx between strain and photoperiod model used to check for significance, (what are the additions called?)
mm1<-aov(lipid_mass~photoperiod*strain+Rep+lean_mass+wet_mass,data=data2)
summary(mm1)
par(mfrow=c(2,2))
plot(mm1)

##The following model removes strain and photoperiod as factors to determine the effect of lean mass and rep on lipid mass
submod<-aov(lipid_mass~Rep+lean_mass,data=data2)

##Adding standardized residuals data to data set
data2$stdresd<-as.vector(scale(resid(submod)))

##Plotting the standardized residuals against strain colored by photoperiod
par(mfrow=c(1,1))
ggplot(data2,aes(x=strain,y=stdresd,colour=factor(photoperiod)))+
  geom_boxplot()

##Strain effect by std residuals of lipid mass
ggplot(data2,aes(x=strain,y=stdresd))+
  geom_boxplot()

##Photoperiod effect by std residuals of lipid mass
ggplot(data2,aes(x=photoperiod,y=stdresd))+
  geom_boxplot()
```


```{r}
#converting wide to long. This code puts all the observations into one column with the label of your choice (measurements used below)
dat.long=gather(data,traits,measurement,wet_mass:lipid_mass)
````

##General Approach
  * Fit a regression model to predict variable (Y).
  * Obtain the predicted and residual values associated with each observation on (Y).
  * Plot the actual and predicted values of (Y) so that they are distinguishable, but connected.
  * Use the residuals to make an aesthetic adjustment (e.g. red colour when residual in very high) to highlight points which are poorly predicted by the model.
 
Simple Linear Regression

#1. Step 1: fit the model: First, we will fit our model. In this instance, let’s copy the dataset to a new object d so we can manipulate it later
````{r}
data1=subset(data,data$treat!="BE12")
data2=subset(data1,data1$treat!="UZ12")
data3=subset(data,data$treat!="BE16")
data4=subset(data3,data3$treat!="UZ16")


d=data
fit=lm(d$lipid_mass ~ d$treat, na.action = na.exclude)
summary(fit)  # Report the results
par(mfrow = c(2, 2))  #Split the plotting panel 
plot(fit)  # Plot the model information
````

#Step 2: obtain predicted and residual values: Next, we want to get predicted and residual values to add supplementary information to this graph. We can do this as follows:
````{r}
d$predicted=predict(fit)   # Save the predicted values
d$residual=residuals(fit) # Save the residual values
````

#Step 3: plot the actual and predicted values: Plotting these values takes a couple of intermediate steps. First, we plot our actual data as follows:
````{r}

ggplot(d, aes(x = treat, y = lipid_mass)+  # and Plot the actual points
  geom_point()+ # Set up canvas with outcome variable on y-axis
  geom_point(aes(y = predicted), shape = 1)+ # Add the predicted values
  geom_segment(aes(xend = treat, yend = predicted))+
  geom_point(aes(y = predicted), shape = 1))
````


````{r}
## here are the plots
## all the trait measurements against treat as boxplot
ggplot(dat.long,aes(x=treat,y=measurement))+
  geom_boxplot()+
  facet_grid(traits~.,scales="free")

## all the trait measurements against treat as boxplot
ggplot(dat.long,aes(x=treat,y=measurement))+
  geom_boxplot()+
  facet_grid(traits~.,scales="free")
````

````{r}
## The response of each trait within each treatment (read:strain and photoperiod)
ggplot(dat.long,aes(x=treat,y=measurement,colour=treat))+
  theme_classic()+
  geom_boxplot()+
  ggtitle("Trait Measurements")+
  scale_colour_discrete(name="Strain x PP")+
  facet_grid(traits~.,scales="free")
````

##Exploring the stats using ANOVA
````{r}
##Changing the number of panels on each output
par(mfrow = c(2, 2))

###Subsetting data into UZ and BE sets for separate anova analysis
data1=subset(data,data$treat!="BE12")
data2=subset(data1,data1$treat!="BE16")
data3=subset(data,data$treat!="BE16")
data4=subset(data3,data3$treat!="BE16")

##constructing an aov model for UZ where "modaUZ" is additive and "modbUZ" is multiplicative
moda=aov(lipid_mass~treat*Rep*lean_mass,data=data)
summary(moda)
plot(moda)

modb=aov(lipid_mass~treat*Rep*lean_mass,data=data)
summary(modb)
plot(modbUZ)

##constructing an aov model for BE where "modaBE" is additive and "modbBE" is multiplicative
modaBE=aov(lipid_mass~treat+Rep+lean_mass,data=data4)
summary(modaBE)
plot(modaBE)

modbBE=aov(lipid_mass~treat*Rep*lean_mass,data=data4)
summary(modbBE)
plot(modbBE)
`````

##Exploring the Stats using AIC
`````{r}
## checking model with AIC
mod1=stepAIC(moda, direction="both")
summary(mod1)

mod2=stepAIC(modbUZ, direction="both")
summary(mod2)
head(data)

## checking model visualy : BE strain
mod3=stepAIC(modaBE, direction="both")
summary(mod3)





mod4=stepAIC(modbBE, direction="both")
summary(mod4)

````

###Calculating the means. Visualized using boxplots and point plots 
````{r}


mean.dat=ddply(data, c("treat","lipid_mass"),summarize,
               N    = length(lipid_mass),
               mean = mean(lipid_mass),
               sd   = sd(lipid_mass),
               se   = sd / sqrt(N))
summary(mean.dat)

ggplot(data,aes(x=treat,y=lipid_mass))+
  geom_boxplot()+
  geom_line(data=mean.dat,aes(x=treat,y=mean))

ggplot(data,aes(x=treat,y=lipid_mass,colour=Rep)+
         geom_point()+
         geom_line(data=mean.dat,aes(x=treat,y=mean,colour=Rep))

data$resid=scale(resid(aov(modaUZ)))
mean.dat2=ddply(data,.(treat,Rep),summarize,mean=mean(data$resid),sd=sd(data$resid))

ggplot(data,aes(x=treat,y=resid,colour=Rep))+geom_boxplot()+geom_line(data=mean.dat,aes(x=treat,y=mean,colour=Rep))

ggplot(mean.dat,aes(x=treat,y=mean,colour=Rep,group=Rep))+geom_line(position="identity",size=2)+geom_errorbar(aes(ymin=mean-sd,ymax=mean+sd),width=.35,size=2)+geom_point(size=10)+ylab("Standardized Residuals of Lipid Mass")
`````

````{r}
##subsetting out extreme sample lines in wet larvae 
#subset(data,Lipid.Wt>0.63)

## removing data lines via subset
#sub.out=subset(data,WET.Larva<0.6)

## or 

#sub.out1=data[-95,]
#dim(sub.out)